---
title: "Assignment 1"
author: "Mary Smith"
date: "Sunday, July 19, 2015"
output: html_document
---


In this assignment we need to process activity data.

##Loading and preprocessing the data

The first thing is to load the data. I store it into a variable called `set`. 

```{r}
set <- read.csv("activity.csv")
head(set)
```

##What is mean total number of steps taken per day?

Calculating the total number of steps taken per day using the tapply function. To avoid na's I used `na.rm = TRUE` parameter.

```{r}
daySteps <- tapply(set$steps, set$date, sum)
```

Making a histogram.

```{r}
hist(daySteps, col = "light blue", main = paste("Histogram of" , "steps per day"), xlab = "number of steps")
```

Calculating the mean and the median of the total number of steps.

```{r}
mean(daySteps, na.rm = TRUE)
median(daySteps, na.rm = TRUE)
```

##What is the average daily activity pattern?

Getting average number of steps.

```{r}
AvSteps <- aggregate(steps ~ interval, set, mean)
```

Making a time series plot of the 5-minute interval.

```{r}
plot(AvSteps$interval, AvSteps$steps, type = 'l', col = "dark green", main = "Daily activity pattern", xlab = "Interval", ylab = "Number of steps" )
```

The maximim number of steps is in the 835 interval.

```{r}
subset(AvSteps, steps == max(AvSteps$steps))
```

##Imputing missing values
The only column which has na's is `steps`. So there are 2304 rows with unknown values.

```{r}
summary(set)
```

Replacing na's with the mean for corresponding day using the `AvSteps`.

```{r}
for (i in 1:nrow(set)){
        if (is.na(set$steps[i])){
                val <- set$interval[i]
                id <- which(AvSteps$interval == val)
                val <- AvSteps$steps[id]
                set$steps[i] <- val
                }
        }
```

Making a histogram.

```{r}
hist_set <- tapply(set$steps, set$date, sum)
hist(hist_set, col = "gray", main = paste("Histogram of" , "steps per day"), xlab = "number of steps")
```

Calculating the mean and the median. There is no change in mean value comparing to original dataset and unsignificant change in median.

```{r}
mean(hist_set)
median(hist_set)
```

##Are there differences in activity patterns between weekdays and weekends?

Creating a new factor variable to indicate type of the day

```{r}
Sys.setlocale("LC_TIME", "English")
set$date <- as.Date(strptime(set$date, format="%Y-%m-%d"))
days <- weekdays(set$date)
set$day_type <- ifelse(days == "Saturday" | days == "Sunday", "Weekend","Weekday")
AvSteps <- aggregate(set$steps, by = list(set$interval, set$day_type),mean)
names(AvSteps) <- c("interval", "day_type", "steps")
```

Making a panel plot containing a time series plot of the 5-minute interval.

```{r}
library(lattice)
xyplot(steps ~ interval | day_type, AvSteps, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```




